
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

     D O    N O T    E D I T      I N      G O O G L E

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\section{MOOCs in a Changing Residential Education Landscape}

% editor: Armando Fox
% participants: Fran\c{c}ois Bry, Carlos Delgado Kloos, Jens Dittrich,
% Armando Fox, Martin Wirsing, Armin Weinberger


We are experiencing a paradigm shift in education. Paradigm shifts have
occurred in other sectors, like the news or the music sector. When this
happens, an unbundling and bundling of components is necessary. In the
case of higher education, and in particular undergraduate education,
certain modes of interaction, certain roles and certain structures have
been established for a long time. It is necessary to think outside the
box to identify new ones that take full advantage of educational
technology, that consider how students make use of technology, that take
into account that the sources of knowledge have changed and that it is
needed to teach the new skills needed in the 21st century in an
effective way. MOOCs are in a way taking the role of books as sources of
knowledge. But the content has to be adapted to the local circumstances,
the place in the curriculum, the prior knowledge, etc.

With the wide availability of high-quality, (often) free, and
easy-to-reuse content online, the instructors' primary responsibility
shifts \emph{from creating content to creating context.} 
In a residential setting,
this creates both disruptions and opportunities that require rethinking
the traditional organization of lectures, labs, recitations, and so on, as
well as the traditional roles and responsibilities of teaching staff:

\begin{itemize}

\item Instructors' responsibilities will shift from creating content to
  creating context for student learning.

\item Rearrangements of residential course elements (flipped classrooms
  being one example) and new teaching roles beyond simply ``instructor''
  and ``teaching assistant'' will better fit this new instructional
  modality.  One cost of this reorganization is changes to campus
  infrastructure, but a benefit is that inexpensive video production
  will empower domain experts to create 
  better materials to prepare and train the new teaching roles.

\item The absence of the ``sage on the stage'' will open new ways to
foster teacher and learner commitment.  Campuses should
focus less on conveying content-oriented skills and more on
social/professional skills, such as collaborative work and
perspective-broadening activities, to complement  independent study
and discovery.

\end{itemize} 

We conclude that modern universities must embrace this disruptigve
technology ecosystem as vigorously as European Renaissance universities
embraced mechanized printing to enhance and cement the
visibility of their intellectual leadership.


\subsection{Instructor responsibilities: Context vs Content}

The abundance of learning material online---instructional videos,
slides, discussion forums, open educational resources,
Wikipedia---challenges the traditional role of teachers and lecturers as
content creators, a role rooted in times when such materials were not
easily available.
Today, lecturer-authored content competes with hundreds of alternatives,
some better and some worse than his own, and the lecturer
loses the content monopoly he has had since classical times.

Students in a traditional lecture first need to complete the first two
learning steps in Bloom's taxonomy: remember and understand the content
presented by the lecturer.
Today, MOOC-driven learning materials used in blended teaching formats
like Flipped or Inverted Classroom provide greater opportunity for
self-study.
In this scenario, the lecturer's role focuses on contextualizing
existing content.
Numerous metaphors have been used to describe this role: weaving a
narrative through existing content, connecting the dots, acting as an
expert tour guide, or as in Vannevar Bush's prescient Memex vision,
blazing a trail.
The lecturer presents his or her personal view on the content based on
domain expertise, puts it into perspective with examples, applications,
analogies or anything else he/she deems useful. The lecturer questions
the content's assumptions and perspectives, helping students evaluate
alternative explanations of content and thereby guiding them in applying
the material (the third step in Bloom’s taxonomy).

The lecturer therefore mediates between  the students and
a ``tsunami'' of unconnected content, selecting 
and remixing it into  a coherent story analogously to how
a film editor reduces hundreds of hours of film footage into a coherent
90-minute movie.
This potentially brings the lecturer's role into direct conflict
with restrictions on remixing imposed by copyright issues, as
discussed in Right to Remix (http://right2remix.org) and other

The role of lecturers as guides applies across disciplines as well as within a specialty.  For example, ``building a user-friendly and secure
database''  crosses at least three different Computer Science specialties
(databases, human-computer interaction, and security). 
This kind of interdisciplinary
story-weaving is facilitated by the abundance of online material which
could be recomposed and selected for this kind of course---perhaps even
without requiring three different professors to teach it.

Interestingly, this instructional format---instructor guidance around
inquiry-based use of existing materials---is the current model for
graduate research.  While undergraduates would be exploring an existing
body of knowledge rather than discovering new knowledge, it is possible
that the undergraduate educational process could become more like
research and less like the unidirectional presentation of information
that dominates undergraduate education today.


\subsection{Course reorganization, new teaching roles,  and the impact of
  inexpensive content creation} 

(LArge courses - copy text from Armando's position)

%% TBD course reorg
%% TBD You can spend a lot of computing power per student; Teacher support is needed for some tasks; how is this going to evolve in the long run?



At the other extreme, MOOCs could provide an opportunity to save small
``boutique'' courses and curricula.
In Europe, after the Bologna Declaration, highly specialized courses and
curricula such as Indian Studies or Albanology had to be integrated into
more general Bachelors' or Masters' curricula such as Asian Studies or
Mediterranean Studies, threatening the existence of important building
blocks for certain cultures.
Several professors around the world could collaboratively produce a
series of specialized MOOCs that would form the basis of ``boutique''
degree programs that would enable continued deep scholarship in such
areas despite limited resources at any single university.

%% transition


Video production and distribution are now inexpensive enough that
individual instructors can create these assets on their own, rather than
requiring hand-holding from a production team. Similarly, computing
power is now inexpensive enough to support compute-intensive interactive
activities even for large numbers of people, such as sophisticated
automated assessment. While various MOOCs have demonstrated the benefits
of these developments for students, the developments are relevant to
classroom instruction in a different but equally important way: they
facilitate the training of new kinds of teaching staff, potentially
allowing us to serve more students with a sublinear increase in faculty.

For example, some skills are better demonstrated by video than explained
in text. Screencasts are very popular in computer science for showing
how to work with particular tools by ``looking over the shoulder'' of an
expert who narrates as he works. Physical skills, such as conducting an
orchestra or working with an artistic medium, can be readily
demonstrated on video much more quickly and effectively than they could
be described textually. Inexpensive video creation and distribution
means that the domain expert can now be the person creating the
materials, eliminating the ``translation gap'' that might arise when
working with a video producer who is not a domain expert.

A classroom instructor can exploit this opportunity to quickly and
easily create training materials \emph{for other instructors} that are
richer and more interactive than a textbook. For example, an instructor
might create a SPOC (Small Private Online Course) targeted not at
students but at other instructors or at teaching assistants.  These
materials would help train teaching staff to assist students in specific
ways: familiarizing staff with a particular homework assignment and
common student problems, giving staff suggestions for reviewing material
before an exam, giving them advice on how to handle problematic
situations in the classroom, and so on.  By exploiting the ability to
create such materials, it becomes possible to train new strata of
teaching staff that further leverage the effectiveness of the lead
instructor, potentially allowing us to educate more students with a
sublinear increase in instructor resources.



\subsection{Teacher and learner commitment in social learning}



Recent meta-analyses (Hattie, 2009) have examind what makes teaching
effective and how the best teachers activate learners with feedback,
challenging goals, direct instruction, and frequent testing, acting as
behavioral organizers rather than facilitating learners to digest
learning material and fulfill tasks.
Just as the 2-sigma finding of tutorial instruction has inspired
intelligent cognitive tutors, these meta-analyses may inspire the
improvement of MOOCs or
whatever follows them.

An inherent challenge of MOOCs is
that learners cannot easily be personally involved, committed, and
activated to a productive learning trajectory. This may not only be a
problem of lacking meta-cognitive skills for self-regulation, but
of different cultural understandings of what education and  courses are, as
well as the model of learning and instruction that is inherent to many
MOOCs, i.e. transmission of information bits with a teacher who may be
both geographically distant and socially disconnected from her students.

To address this challenge, educational technology
approaches may be: provide for introductions that include compressed
assessment, self-positioning/introductions and trainings in
meta-cognitive strategies for online learning of participants develop
compiled forms of automated and peer assessment that can be used
continuously and fed back to students instantly, but also aim for
summative evaluation provide computer-supported, ready-to-use scripts
for teachers and learners that build on a robust body of evidence, e.g.,
reciprocal teaching, direct instruction, mastery learning provide roles
to students and teachers adaptively orchestrate learning activities in
different learning arrangements from individual to small group to
community interactions including dynamic regrouping of students arrive
at new rhythms for learning (breaks, inverted classrooms, independent
study time, micro learning, time for argumentation) cater for
connections from MOOC to local courses to online presences of students

 Should campuses capitalize on their social and professional
  networking benefits, teaching skills that are less content-oriented
  and more crosscutting such as teamwork and collaboration and focusing
  on placing students in internships or international experiences that
  will broaden their perspectives?

 Campus infrastructure today is highly optimized around traditional
  delivery models.  What is
the role of personal devices such as smartphones and tablets for social
learning interaction? Should campuses 
  have fewer computer labs and more open space for collaboration?  


\subsection{Summary: MOOCs and the intellectual prominence of universities}

The invention of the printing press led to a  ``technical-intellectual
ecosystem'' around printing consisting of print shops, skilled laborers,
authors, editors, readers, publishers, and investors---the ''venture
capitalists'' who recognized the 
disruptive potential of the new medium both to turn a profit and to
spread ideas they 
believed in~\cite{lefebvre}.
These ecosystems increased the visibility not only of previously-obscure
founding texts of humanism, but also of some universities.
Those universities surrounded by such ecosystems---Frankfurt, Paris,
Strasbourg, Leiden, Venice---became the most influential of their
time, as the ideas and writings of their faculty gained wide visibility.

The same is now happening with MOOCs and the forthcoming
new educational media which some universities, mostly in North America,
are developing and with which they are experimenting. 
This new
educational medium is not only an opportunity to make a profit, but also
to spread ideas and gain reputation. They are an
essential vector of success and visibility in today's higher
education: they will help attract the best students and
teachers, provide them with up-to-date working environments, and in doing
so, contribute to their success and to the success of their educational
institutions.  Universities wishing to retain their intellectual
prominence ignore this technology at their peril.



